---
title: "Step01_SetupEnvironment"
author: "Max Bennett"
date: "2025-06-06"
output: html_document
---

While KrSPLAT has been put together as something of an 'a la carte' resource that can be easily copy-pasted from, in order to improve ease of use for the resource as a whole, much of the code requires packages, helper functions, and variables established in this first document. To download helper functions, go to the [kroganlab github](https://github.com/kroganlab)

# Project Settings

Please review the names and paths in the following chunk carefully, they should all be specific to your project!

```{r select local directories and naming}

projectName <- "<replace with your project name>" # pick something short and recognizable, 
                                                  # note that all saved files will have the datetime appended to the name, so you don't have to include that here.

searchTool <- "spectronaut" # One of c("spectronaut", "maxquant", "fragpipe", "diann")

experimentalVariables <- c("treatment") # list of experimental variable names, ex c("drug", "celltype", "infection"), c("drug", "taggedProtein")
                                        # if more than one experimental variable is listed, Condition names in the input data are expected to be separated by underscores, in order
                                        # ex. experimentalVariables = c("drug", "celltype", "infection")
                                        # ex. Conditions = c("veh_HEK_uninf", "veh_HEK_uninf", "drug_HEK_inf", "drug_Macro_inf", etc. )

inputDataFolder <- "./" # Filepath your input data is located in.
                        # Default is the same directory as this .Rmd
  
outputDataFolder <- ""  # Filepath you want to save intermediate and final files to. This is for data and images.
                        # Default is a blank string "", which will be automatically set to "<nameOfYourRmd>_data"

autoSaveFigs <- T     # Default setting to save all generated figures (or not), this can be toggled at any point by changing the assigned value

preferredFormat <- "pdf" # one of c("png", "pdf", "both")

#' Ensure these filepaths to the bioinformatic team's helper functions are correct...
#' If you do not already have them downloaded, they can be found on the kroganlab github (linked above). 
bpUtilsPath <- "~/Documents/Code/bp_utils/"
mbUtilsPath <- "~/Documents/Code/maxb_utils/"
yzUtilsPath <- "~/Documents/Code/Yuan_utils/"
# mgUtilsPath <- "./"   # Not in use currently
  
aestheticParameterFile <- "defaultAesthetics.yaml" # this is a developing feature, leave as default for now

```

# Automated Setup

## Setup .Rmd Settings 

```{r markdown options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup Package Libraries

```{r load generic package librarires, include=FALSE}
suppressPackageStartupMessages({
library (data.table) # powerful syntax for data wrangling
library (pbapply) # implements _apply functions with progress bars and parallelization
library (rstudioapi) # 
library (devtools)  #
library (yaml)      #
library (readxl) # reads excel files
  })
```

```{r load packages for data visualization, include=FALSE}
suppressPackageStartupMessages({
library (ggplot2) # modular plots
library (ComplexHeatmap) # plots heatmaps
library (viridis) # continuous color palettes 
library (scico)  # discrete color palettes
library (ggrepel) # adds force-directed labels to ggplot
library (ggpubr) # adds features from graphpad prism to ggplot
library (ggridges) # adds joyplots to ggplot
library (pROC) # adds simple roc curve plot to ggplot 
library (circlize) # adds circos plots and colorRamp2
  })
```

```{r load packages for bioinformatics, include=FALSE}
suppressPackageStartupMessages({
library (MSstats)
library (clusterProfiler)
library (fgsea)
library (cRomppass)
  })
```

## Setup Local Environment

```{r source helpers and local parameters, include=FALSE}
source(file.path(bpUtilsPath, "ManageScriptData.R"))
source(file.path(mbUtilsPath,"GenericHelpers.R"))

```

```{r create directories, include=FALSE}
cat("Current directory is ", getwd())
if ( !file.exists(paste0(getwd(),"/logs")) ){
  cat("/logs directory does not exist--making /logs directory.")
  dir.create(paste0(getwd(),"/logs"))     }
dataFolder <- ScriptNamedDir()
calledFrom <- substrLeft( tail(strsplit(dataFolder,"/") [[1]], n = 1), 0, -5 )
cat("Logging installed package versions.")
WriteSessionInfo(path = paste0( "./logs/", DateFileName(paste0(calledFrom, "_SessionInfo.txt")) ) )
WriteInstalledPackages(path = paste0( "./logs/", DateFileName(paste0(calledFrom, "_InstalledPackages.txt")) ) )
```
